Title here

### 2020-09-01 analysis/paper/paper.tex
mas draft

@@

We used monthly geopotential height at 2.5 longitude by 2.5 latitude
resolution from ERA5 \citep{hersbach} for the period 1979 to 2018.

@@

@@

These results point to very different behaviours between the
stratospheric and tropospheric SAM. \#FIXME

This suggests that both the Asymmetric and the Symemtric component of
the tropospheric SAM are highly vertically coherent, both in their
individual evolution and their temporal relationship. This is to be
expected since the SAM is mostly equivalent barotropic (citaaaa).

@@

Interestingly, the structures shown in Figure
\ref{fig:vertical-regression} are surprisignly robust to the choice of
index level. For any stratospheric (above 100 hPa) index, the resulting
anomalies are very similar to the wave-1 structure with maximum near 10
hPa in panel a. Conversely, for any tropospheric (below 100 hPa) index,
the result is very similar to panel b. The pattern cross-correlation
between levels of each segment of the atmosphere is greater than 0.9
(Figure A\#fixme). The patterns mainly change in amplitude. The
tropospheric pattern is maximised by the 300 hPa Asymmetric SAM index
and the stratospheric pattern increased monotonically with height.

@@

In South America (Figure \ref{fig:pp-regr-america}) , the annual-level
regression shows that the SAM is associated with statistically
significant precipitation decreases both in Southeastern South America
and Southern Chile (panel a.1). Both signals were observed by
\citet{silvestri2009} and the former also by \citet{gillett2006}. Panels
b.1 and c.1 show a remarkable clean separation between the Asymmeric SAM
--associated with the signal located in Southestern South America-- and
the Symmetric SAM --associated with the signal in Chile. This separation
is consistent with the mechanisms responsible for these effects. In
Southeastern South America, anomalous meridional winds lead to less
precipitation by inhibiting moisture convergence from the South American
Low Level Jet \citep{silvestri2009}. In Southern Chile, the reduced
westerly flow reduce moisture transport from the Pacific Ocean (cita??).
The is a small area of increased precipitation with SAM near central
Argentina which is also present in the station-based analysis by
\citet{gillett2006} and that is explained by the Asymmmetric SAM.

The negative relationship between precipiation and SAM index in
Southeastern South America and Southern Chile is evident in all seasons
except Winter (although not always stastitically significant) and it
appears to maximise in Autumn.

@@

\citet{silvestri2009} showed that impacts linked to the SAM changed
rather dramatically before and after 1980, and that this change can be
explained by changes in the regional-scale structure of the SAM-related
circulation. In particular, the decreased in precipitation in South
America (Figure \ref{fig:pp-regr-america} panel a.1) was absent or even
switched sign.

\citet{silvestri2009} also thowed

Therefore, is very likely that if we were to repeat this analysis using
pre-satellite data, the resulting Asymmetric SAM would look very
differnt.

This suggests that both the Asymmetric and the Symemtric component of
the tropospheric SAM are highly vertically coherent, both in their
individual evolution and their temporal relationship. This is to be
expected since the SAM is mostly equivalent barotropic (citaaaa).

\citet{hosking2013} observed a pattern similar to the one near the
Amundsen Sea in panel a.4 associated with the strengthening of the ASL.
The positive anomalies of SIC near the Amundsen sea being inconsistent
with the anomalies of meridional wind associated with the ASL. Row 4 in
Figure \ref{fig:regr-ice} might shed some light to resolve this mistery.
The SAM is evidently related to the ASL through it's Asymmetric
component, and since the ASymmetric SAM is correlated with the Symmetric
SAM, the ALS will be rleated to the symemtric SAM. Consiering the
similarities between the the SIC anomalies observed by
\citet{hosking2013} (in its Figure 7) and Figure \ref{fig:regr-ice}
panel a.4, it is very likely that what was observed in that paper was a
residual effect of the SAM. \#FIXME!!

@@

\begin{figure}
\includegraphics{A6-1} \appendcaption{A6}{Regressions maps resulting from performing one multiple regression (column a) and from performing two simple regressions (column b)}\label{fig:A6}
\end{figure}

\begin{figure}
\includegraphics{A7-1} \appendcaption{A7}{Zonal waves derives from the regression maps from performing one multiple regression (column a) and from performing two simple regressions (column b)}\label{fig:A7}
\end{figure}

@@

>>>>

\citet{fogt2012} studied the characteristics of the asymmetric structure
of the SAM. It computed the zonally anomalous component of mean sea
level pressure (MSLP) composites for positive and negative SAM events
and created two indices by projecting MSLP fields ontno them. However,
the use of composites leads to some issues that affects the
interpretability of the results. First, they can be dependent on the
choice of threshold used to define positive and negative events.
Secondly, by discarting data that don't meet the threshold, they don't
use all the information avaiable. Due to the realtive short timeframe
used, this leads to some composites being composed of as little as 4
years. Third, the resulting composites corresponding to each polarity
and season are derived from the average of different amount of fields
and from different years. This last issue is particularly important in
light of the changing structure of the SAM before and after 1980
\citep{silvestri2009}. In \citet{fogt2012}, the DJF SAM+ composite uses
only 7 years, 5 of which are later than 1988, whereas all of the 8 years
used for their DJF SAM- composite are from earlier than 1988.

@@

@@

We used monthly geopotential height at 2.5 longitude by 2.5 latitude
resolution from ERA5 \citep{hersbach2020} for the period 1979 to 2018.

@@

Figure \ref{fig:r-squared-trend} shows decadal trends for the explained
variance of each index. There is no evidencie of a significant trend in
the stratosphere. In the troposphere, there is a positive trend for the
Asymmetric SAM and no significant trend for the Symmetric SAM. This
suggest that the SAM has become more asymmetric in the period from 1979
to 2018. The change is slight, though; of the order of 1\% icreased
explained variance per decade.

\begin{figure}
\includegraphics{r-squared-trend-1} \caption[Decadal trends for explained variance of each index at each level for the period 1979-2018]{Decadal trends for explained variance of each index at each level for the period 1979-2018. Shading indicates the 95\% confidence interval.}\label{fig:r-squared-trend}
\end{figure}

@@

Interestingly, the structures shown in Figure
\ref{fig:vertical-regression} are surprisignly robust to the choice of
index level. For any stratospheric (above 100 hPa) index, the resulting
anomalies are very similar to the wave-1 structure with maximum near 10
hPa in panel a. Conversely, for any tropospheric (below 100 hPa) index,
the result is very similar to panel b. The pattern cross-correlation
between levels of each segment of the atmosphere is greater than 0.9
(Figure A8). The patterns mainly change in amplitude. The tropospheric
pattern is maximised by the 300 hPa Asymmetric SAM index and the
stratospheric pattern increased monotonically with height.

@@

In South America (Figure \ref{fig:pp-regr-america}), the annual-level
regression shows that the SAM is associated with statistically
significant precipitation decrease in Southeastern South America (SESA)
and Southern Chile and non-significant increase in South Brazil, near
the South Atlantic Convergence Zone (SACZ) (panel a.1).

Panels b.1 and c.1 show a remarkably clean separation between the
Asymmeric SAM --associated with the Southeastern South American and
Southern Brazilian signals-- and the Symmetric SAM --associated with the
signal in Southern Chile. This separation is consistent with the
mechanisms responsible for these effects. In Southern Chile, the reduced
westerly flow reduce moisture transport from the Pacific Ocean (cita??
\#FIXME).

In Southeastern South America, anomalous meridional winds lead to less
precipitation by inhibiting moisture convergence from the South American
Low Level Jet \citep{silvestri2009}. The increased precipiation in the
South Atlantic Convergence Zone, on the other hand, appears to be
related to modulation of the anomalous SAM circulation in SACZ events
which lead to more frequent and intense SACZ events during positive SAM
\citep{rosso2018}.

There is a small area of increased precipitation with SAM near central
Argentina which is also present in the station-based analysis by
\citet{gillett2006} and that is explained by the Asymmmetric SAM.

Except during Winter, the seasonal-level regression all show these
similar patterns although not as cleanly and only in some cases
statistically significant.

@@

We presented a method to systematically separate the zonally asymmetric
and zonally symmetric components of the SAM. By itself, the spatial
structure and temporal evolution of the SAM shows a strong separation
between the stratosphere SAMs. The first EOF in the stratosphere is
monopolar in nature while the first EOF in the troposphere is a proper
annular mode. Their repespective departures from the zonal mean are also
different: a zonal wave 1 dominates the stratospheric SAM, waves 3 and 2
dominate the tropospheric SAM. Furthermore, there is little temporal
correlation between the tropospheric and stratospheric time series.

The zonal asymmetric component of the SAM at each level is even more
decupled between the troposphere and the stratosphere. Their temporal
evolution shows essentialy zero correlation (Figure
\ref{fig:cross-correlation}) and the signal associated with the
stratospheric Asymmetric SAM is completely restricted to the
stratosphere (Figure \ref{fig:vertical-regression} panel a).
Geopotential height anomalies associated with the tropospheric
Asymmetric SAM, on the other hand, do extend to the stratosphere, but
those anomalies do not project strongly into the stratospheric
Asymmetric SAM.

We show that the observed positive trends towards positive SAM is
restricted to the tropospheric SAM and is explained by the Symmetric
component (Figure \textbackslash{}ref\{fig:trends{]}). However, the
degree of asymmetry appears to have increased slightly in the last 40
years, as the Asymmetric SAM explains an increasingly proportion of the
variance (Figure \ref{fig:r-squared-trend}).

In terms of impacts, we see that ..Temperature\ldots{} \#FIXME

The patterns of SAM-associated precipitation anomalies are similarly
well separated. In South America, we show that negative anomalies
observed in Chile related to the SAM are well explained by the Symmetric
compomnet, while the precipitation dipole in Southern South America and
the South Atlantic Convergence Zone is explained by the Asymmetric
component.

\subsubsection{Limitations}

Our method assumes linearity in the asymmetric component of the SAM.
That is, assumes that zonal symmetries associated with positive SAM are
oposite and equal to the ones associated with negatie SAM.
\citet{fogt2012}'s composites suggest that this might not be entirely
valid, although we argue that much of that apparent non-linearity is due
to the heterogenous nature of the selected years for constructing the
composites. Using our data (from 1979 to 2018), seasonal composites of
zonal anomalies of 700 hPa geopotential height for for SAM+ and SAM-
show pattern linear correlations greater than -0.7 for all seasons and
are visually very linear (Figure A9). Therefore, we belive that our
method is at the very least a reasonable approximation of the fenomenon.

We also asumed that the structure of the SAM zonal anomalies is stable
in all seasons. Again, this is not unreasonable, as geopotential zonal
anomalies computed by projecting the first EOF \emph{of each season} are
very similar to each other (Figure A10).

\citet{silvestri2009} showed that impacts linked to the SAM changed
rather dramatically before and after 1980. In particular, the negative
relationship with precipitation in South America (consistent with Figure
\ref{fig:pp-regr-america} panel a.1) was absent in some areas and
switched sign in other in the earlier period. The correlation between
ENSO and SAM is similarly non-stationary, also disapearing before 1973.

Seeing as both the ENSO-SAM relationship and most of the precipitation
imacts in South America are captured by the Asymmetric SAM, the results
presented here are most likely period-dependent. Therefore, is very
likely that if we were to repeat this analysis using pre-satellite data,
the resulting Asymmetric SAM would look very different.

@@

@@

\begin{figure}
\includegraphics{A9-1} \caption[700 hPa Geopotetnial height zonal anomalies of composites of positive and negative SAM months selected using 1 standard deviation as threshhold]{700 hPa Geopotetnial height zonal anomalies of composites of positive and negative SAM months selected using 1 standard deviation as threshhold. Numbers in the column headers are pattern correlation between SAM+ and SAM- composites and number of monthly fields used to construct the composites.}\label{fig:A9}
\end{figure}

\begin{figure}
\includegraphics{A10-1} \caption[Zonal of projection of 700 hPa onto the first EOF of each season]{Zonal of projection of 700 hPa onto the first EOF of each season.}\label{fig:A10}
\end{figure}
